\documentclass{article}
\usepackage{amsmath}


\begin{document}
	
\section*{Two-Way FE Difference in Difference}

By construction, the difference-in-difference estimate of the average treatment effect of $X_{it}$ on $Y_{it}$ in a linear equation is $E[Y_{21} - Y_{11}]$ (difference between treated/control at $t=1$) minus $E[Y_{22} - Y_{12}]$ (difference between treated/control at $t=1$). In the two-way FE model of D-in-D, $\beta$ serves as an estimand for the ATE.

Using the following linear models:

\begin{align*}
	Y_{11} &= \alpha + \beta * X_{11} + \eta_1 + \nu_1 \\
	Y_{12} &= \alpha + \beta * X_{12} + \eta_1 + \nu_2 \\
	Y_{21} &= \alpha + \beta * X_{21} + \eta_2 + \nu_1 \\
	Y_{22} &= \alpha + \beta * X_{22} + \eta_2 + \nu_2 
 \end{align*}
 
 and a binary treatment $X$ with $X=1$ if $i=2$ and $X=0$ if $i=1$. (The treatment variable and the units are perfectly collinear). 
 
 Then, the first difference, or the artificial control, is equal to $E[Y_{21} - Y_{11}]$, which from the equations above reduces to:
 
 \begin{align*}
 	E[Y_{21} - Y_{11}] = \beta * X_{21} + \eta_2 - \beta * X_{11} - \eta_1 
 \end{align*}

because the time fixed effects will cancel each other at this time point. Similarly, the second difference, or the artificial treatment, $E[Y_{22} - Y_{12}]$, is equal to 

\begin{align*}
E[Y_{21} - Y_{11}] = \beta * X_{22} + \eta_2 - \beta * X_{12} - \eta_1 
\end{align*}

Taking the difference of these differences yields the following expression:

\begin{align*}
	E[(Y_{22} - Y_{12}) _(Y_{21} - Y_{11}) )] = \beta*X_{22} - \beta*X_{12} - \beta * X_{21} + \beta * X_{11}
\end{align*}

Because the unit fixed effects cancel out in this difference. However, as written above, if $X=1$ if $i=2$ and $X=0$ if $i=1$, then the expression above reduces to the following:

	\begin{align*}
	E[DD_{2FE}] &= \beta * 1 - \beta * 0 - \beta * 1 + \beta * 0 \\
	E[DD_{2FE}] &= 0
\end{align*}

So, for all possible $\beta$, the true two-way FE estimate is zero. (Or check my math if I did something wrong). That means that any non-zero values will come from bias in the errors (i.e. they are not in fact mean zero), or if the panels are not balanced (all the $\beta$s are not the same).

\section*{Interactive Between Effect D-in-D}

If, on the other hand, we use the interactive between effects estimator, then we are interested in the effect $\delta$ as an estimand for the ATE. As above, $X=0$ if $i=1$ and $X=1$ if $i=2$, and the equations are as follows:

\begin{align*}
Y_{11} &= \alpha + \beta * X_{11} + \nu_1 + \delta * \nu_1 * X_{11} \\
Y_{12} &= \alpha + \beta * X_{12} + \nu_2 + \delta * \nu_2 * X_{12}\\
Y_{21} &= \alpha + \beta * X_{21} + \nu_1 + \delta * \nu_1 * X_{21}\\
Y_{22} &= \alpha + \beta * X_{22} + \nu_2 + \delta * \nu_2 * X_{22}
\end{align*}

Then, the first difference, or the artificial control, is equal to $E[Y_{21} - Y_{11}]$, which from the equations above reduces to:

\begin{align*}
E[Y_{21} - Y_{11}] = \beta * X_{21} + \delta * \nu_1 * X_{21} - \beta * X_{11} - \delta * \nu_1 * X_{11}
\end{align*}

Because the time fixed effects will cancel each other at this time point. Similarly, for the second difference, or the artificial treatment, $E[Y_{22} - Y_{12}]$, is equal to 

\begin{align*}
E[Y_{21} - Y_{11}] = \beta * X_{22} + \delta * \nu_2 * X_{22} - \beta * X_{12} - \delta * \nu_2 * X_{12} 
\end{align*}

Taking the difference of these differences yields the following expression:

\begin{align*}
E[((Y_{22} - Y_{12}) - (Y_{21} - Y_{11}))] =&  \beta*X_{22} + \delta * \nu_2 * X_{22} - \beta*X_{12} - \delta * \nu_2 * X_{12} \\ & -  \beta * X_{21} - \delta * \nu_1 * X_{21} + \beta * X_{11} + \delta * \nu_1 * X_{11}
\end{align*}

To plug in the actual values of the variables $X$ and $\nu$, $X$ is either 0 or 1 as before, and $\nu=0$ for $\nu_1$ and $\nu=1$ for $\nu_2$ because the first time point will serve as a reference category in the regression. This yields the following expression:

\begin{align*}
E[DD_{IBE}] &= \beta * 1 + \delta * 1 * 1 -  \beta * 0 - \delta * 1 * 0\\ &- \beta * 1 - \delta * 0 * 1 + \beta * 0 + \delta * 0 * 0 \\
E[DD_{IBE}] &= \delta
\end{align*}

So the interactive between effect version of the D-in-D estimator does return $\delta$ as an unbiased quantity in expectation, assuming that the errors are mean zero (i.e., the parallel paths assumption).

\end{document}